Open-source landscape for 3D CSEM modelling

### 2020-10-23 manuscript/Manuscript.tex
Before moving to GJI template

% Additional packages
\usepackage[UKenglish]{babel}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{amssymb, amsmath, amsfonts}
\usepackage{siunitx}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{xspace}
\usepackage{flafter}
\usepackage[pdftex, final, allcolors=blue, colorlinks=true]{hyperref}

@@

\begin{abstract}
Large-scale modelling of three-dimensional controlled-source electromagnetic (CSEM) surveys used to be feasible only for large companies and research consortia. This has changed over the last few years, and today there exists a selection of different open-source codes. Using four different codes in the Python ecosystem, we perform simulations for increasingly complex models. We first verify the computed fields with semi-analytical solutions for a simple layered model. Then we validate the responses of a more complex block model by comparing results obtained from each code. Finally we compare the responses of a real world model with results from the industry. These validations show that the open-source codes are able to compute comparable CSEM responses for challenging, large-scale models. Our comparison includes finite-element and finite-volume codes using structured rectilinear and octree meshes as well as unstructured tetrahedral meshes. Accurate responses can be obtained independently of the chosen method and the chosen mesh type. The runtime and memory requirements vary greatly based on the choice of iterative or direct solvers. However, we have found that much more time was spent on designing the mesh and setting up the simulations than running the actual computation. The challenging task is, irrespective of the chosen code, to appropriately discretize the model. We provide three models, each with their corresponding discretization and responses of four codes, which can be used for validation of new and existing codes. The collaboration of four code maintainers trying to achieve the same task brought in the end all four codes a significant step further. This includes improved meshing and interpolation capabilities, resulting in shorter runtimes for the same accuracy.
%
% https://academic.oup.com/DocumentLibrary/GJI/gji%20keyword%20list%20updated2017.pdf
\\[1cm] \textbf{Keywords}: Controlled source electromagnetics (CSEM), Numerical modelling, Electrical properties
\end{abstract}

\newpage
\subsection*{Discussed items re manuscript}
\begin{itemize}
  \item Orthography: BE for GJI; using discretize.
  \item Author order is fine by everyone.
  \item Everyone is responsible that the info of their codes is correct (machine info, runtime, memory, dof, etc) and agrees with the info in the \texttt{*.nc}-files.
  \item The runtime in the table is real-world time. So actual computation time is roughly `nprocs x runtime`; also, nprocs includes processes and threads.
  \item Regarding submission:
    \begin{description}
      \item[Current submission] RR, OCR, LH have no submission with GJI currently.
      \item[History] We will keep all git history.
      \item[Repo] \href{https://github.com/swung-research/3d-csem-open-source-landscape/}{github.com/swung-research/3d-csem-open-source-landscape}.
      \item[Suggested reviewers] We suggest the following reviewers, mainly because they have the relevant experience and it would be interesting for us to have their feedback: \emph{Kerry Key} (domain knowledge as developer of \texttt{MARE3DEM}); \emph{Rune Mittet} (domain knowledge; he will know what SBLwiz actually does); \emph{Dimitry Avdeev} and \emph{Ralph-Uwe Börner} (both did extensive 3D code overview/review papers in 2005 and 2010, respectively).
    \end{description}
\end{itemize}

@@

The open-source landscape in the field of 3D CSEM modelling changed in the last five years quite dramatically, and here we introduce and compare four recent open-source projects. All presented codes are in the Python ecosystem and use either the FV method on structured grids or the FE method on unstructured tetrahedral meshes. Having different codes which use different methods and different meshes is ideal to address the topic of validation. Analytical and semi-analytical solutions only exist for simple halfspace or layered-Earth models, which served mostly to verify new codes. The only objective possibility to ensure the accuracy of solutions beyond these simple models is by comparing results from different modellers. If different discretizations and implementations of Maxwell's equations yield the same result, it gives confidence in their accuracy. \cite{GJI.13.Miensopust} presents a review of two workshops dealing with the validation of MT forward and inversion codes, but we are not aware of any comparable comparison study or benchmark suite for CSEM data. We hope our efforts help to ensure the reliability of simulations for any new or existing, open-source or closed-source 3D CSEM project.

@@

We briefly present our codes in this section. It is, however, beyond the scope of this article to go into every detail of the different modellers, and we refer to their documentations for more details. An overview comparison of the codes is given in Table~\ref{tbl:codecomparison}. All four have in common that they solve Maxwell's equation in its differential form (DE) under the quasi-static or diffusive approximation, hence neglecting displacement currents, which is a common approximation for the low frequency range commonly applied in CSEM. The machines on which the different codes were run are listed in Table~\ref{tbl:machines} together with the responsible operator. A few clarifying words on abbreviations and definitions: For the boundary conditions (BC) in the comparison table we use the abbreviations \emph{PEC} and \emph{PMC}, which stand for perfect electric conductor
% ($\vec{n}\times\vec{E} = 0; \vec{n}\cdot\vec{H} = 0$)
and perfect magnetic conductor%
% ($\vec{n}\cdot\vec{E} = 0; \vec{n}\times\vec{H} = 0$)
, respectively. Another abbreviation used in the tables is \emph{dof} for degree of freedom, which is equivalent to the size of the SLE we are solving. Finally, we use \emph{runtime} as the wall time or elapsed real-world time from start to end of solving the SLE; pre- and post-processing (e.g., mesh generation) is not measured. Note that the actual computation time or CPU time is therefore much higher than the reported runtime for the codes that run in parallel, with an upper limit of \#Procs$\times$Runtime. Memory refers to the maximum memory increase at any point of solving the SLE.

\tabl[bthp]{codecomparison}{Comparison of the four codes under consideration. Note that \emg3d is a solver on its own, while the other codes implement third-package solvers such as \texttt{PETSc} \citep{Preprint.Abhyankar}, \texttt{MUMPS} \citep{SIAM.01.Amestoy}, or \texttt{PARDISO} \citep{FGCS.04.Schenk}.}{
  \centering
  \footnotesize
\begin{tabularx}{\linewidth}{lXXXX}
  \toprule
  %
  & \custem & \emg3d & \petgem & \simpeg  \\
  \midrule
  %
  Website & \href{https://custem.rtfd.io}{custem.rtfd.io}
          & \href{https://empymod.github.io}{empymod.github.io}
          & \href{http://petgem.bsc.es}{petgem.bsc.es}
          & \href{https://docs.simpeg.xyz}{simpeg.xyz} \\
  License & GPL-3.0 & Apache-2.0 & BSD-3-Clause & MIT \\
  Installation & \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{pip}
               & \texttt{pip}; \texttt{conda} \\
  Comp. Dom. & frequency \& time & frequency & frequency & frequency \& time \\
  Method & FE & FV & FE & FV \\
  Mesh & tetrahedral & rectilinear & tetrahedral & recti- \& curvilinear \\
  BC & PEC; PMC & PEC & PEC & PEC; PMC \\
  Solver & \texttt{MUMPS} & \emg3d & \texttt{PETSc}; \texttt{MUMPS} &
           \texttt{PARDISO}; \texttt{MUMPS} \\
  %
  \bottomrule
\end{tabularx}}%
%

@@

There are a few noteworthy points:

\begin{enumerate}
  \item The error levels between the different codes are overall comparable; the NRMSD is generally below a few percent, increasing towards larger offsets.
  \item The NRMSD between the FV codes \emg3d and \simpeg is comparably low except at the boundary. The generally low NRMSD is because these codes use the same rectilinear mesh for modelling. The increasing NRMSD towards the boundary is because \emg3d uses the PEC boundary condition, while \simpeg uses the PMC boundary condition in this example.
  \item The FE codes \custem and \petgem show similar boundary effects, but are influenced by the anomaly response (higher errors on right side and now unsymmetrical error distribution).
  \item The agreement within the same differential equation method (FE or FV) is better than across methods. 
\end{enumerate}

@@

The responses for all six frequencies and all three electric components are shown in Figure~\ref{fig:results-marlim-responses}, both inline and broadside. The published responses are shown in colour with markers for the frequency, and the responses from our codes are shown beneath in grey colours in this overview plot.
%
\plot*{results-marlim-responses}{width=\textwidth}{MR3D comparison between our codes (grey lines) and the published data (coloured lines). The grey lines under the coloured lines are not visible in most areas, meaning that they are very similar. However, there are three notable zones, (1) to (3), which are explained further in the text.}
%
Two data-sets were published, one containing clean responses, and one where realistic noise was added. For the comparison we use the clean data, but we indicate the chosen noise level by the horizontal line at \num{2e-15}\,V/m. It can be seen that the grey lines are not visible for most parts, which means that the data agree well. However, there are a few notable points where this is not the case, and they are marked with numbers in the figure.
\begin{enumerate}
  \item There are some noticeable differences for positive offsets of the broadside $E_y$ and $E_z$ components, which are likely related to the bathymetry. The reason why we cannot reproduce the published results exactly is that the code is not open-source, and we do not exactly know what it does internally. See the discussion around Figure~\ref{fig:results-marlim_survey}.
  \item The three highest frequencies became noisy at large offsets for all of our codes (well below any real-world noise level). The reason why the published responses are not noisy is probably because they were computed in the time domain, and transformed to the frequency domain. This explains their smooth behaviour, without saying anything about their accuracy.
  \item The responses for the $E_y$ component of the inline acquisition line do not agree at all. In the 1D case, the inline $E_y$ component would be zero, and the only response we can measure here comes from 3D effects. These responses are very low, roughly two orders of magnitude lower than the $E_x$ component. The published responses are not stable either, and therefore any of the responses is as bad as the other. Also, tiniest differences in meshing or different interpolation algorithms will have a huge effect here, as well as the boundary condition of the air-water interface of the published results.
\end{enumerate}

It is important to note that note the published results nor any of our results is correct, but the closer they agree with each other the more certain we can be of the outcomes. Figure~\ref{fig:results-marlim_2published} shows the NRMSDs between the published results and our codes for the three strongest components, the inline $E_x$ and the broadside $E_x$ and $E_y$ components, for three frequencies.
%
\plot*{results-marlim_2published}{width=\textwidth}{NRMSDs for all four codes in comparison with the published data, for the inline $E_x$ field and the broadside $E_x$ and $E_y$ fields in the left, middle, and right column, respectively. Shown are the three frequencies 0.125\,Hz, 0.5\,Hz, and 1.0\,Hz in the top, middle, and bottom row, respectively.}
%
A few conclusions can be drawn from this figure: In general the NRMSD is roughly 10\,\% or less. The structured meshes have in some cases a smaller NRMSD, e.g., for negative offsets in the inline and broadside $E_x$ responses, and in other cases the unstructured tetrahedral meshes, e.g., for positive offsets for all broadside $E_y$ responses, and often they have a comparable NRMSD. Some NRMSDs have an interesting step-pattern, which is more pronounced for the structured meshes than for the unstructured meshes; this is related to the annotated point (3) in Figure~\ref{fig:model-marlim}, and will be discussed in more detail with Figure~\ref{fig:results-marlim_survey}.

The importance of the meshing to the result can be seen in Figure~\ref{fig:results-marlim_2ours}, where we compute the NRMSDs between our codes.
%
\plot*{results-marlim_2ours}{width=\textwidth}{NRMSDs, just as in Figure~\ref{fig:results-marlim_2published}, but comparing some of the four codes with each other. \custem and \petgem produce very similar results, which is due to the fact that they use the same mesh. This also shows that the biggest impact in 3D forward modelling comes from the discretization.}
%
The most obvious result in that figure is that \custem and \petgem produce results that are almost identical, their NRMSD is generally below 0.03\,\%. The reason is simple: \custem and \petgem use the exactly same mesh for the computation, the only difference is therefore the solver. This exemplifies that the biggest difference between different codes is the discretization; the solver is important when it comes to runtime and memory, but not for the accuracy of the result. As these two are so similar we compare \emg3d and \simpeg only to \custem in Figure~\ref{fig:results-marlim_2ours}, as the comparisons to \petgem would look the same. But in general the NRMSDs between our codes look similar as in the comparison with the published results, which is due to the different meshes in use.

The actual differences in meshing is shown in the next two figures. Figure~\ref{fig:results-marlim_big} shows on the top row the rectilinear mesh of \emg3d, on the middle row the octree mesh of \simpeg, and on the bottom row the tetrahedral mesh of \custem and \petgem. The actual mesh, the broadside $|E_x|$ field, and the broadside $|E_y|$ field are shown in the left, middle, and right columns, respectively. The extent of the plot includes the entire computational domain for the tetrahedral mesh (in x and z direction); the rectilinear and octree mesh extents are larger than what is shown (rectilinear: $x=327$ to 454\,km, $z=-57$ to 83\,km; octree: $x=288$ to 492\,km, $z=-123$ to 41\,km).
%
\plot*{results-marlim_big}{width=\textwidth}{Meshes and broadside $E_x$ and $E_y$ fields in the left, middle, and right column, respectively, for the rectilinear, octree, and tetrahedral meshes in the top, middle, and bottom row, respectively. The shown dimensions shows the entire computational domain for the tetrahedra mesh, whereas the entire rectilinear and octree meshes are bigger.}
%
This figure is interesting for various points: The differences between the different meshes are very obvious. Particularly the advantage of octree and tetrahedra, which can create large cells outside of the region of interest to push away the boundary, whereas a rectilinear grid, even though stretched, has to compute many cells far away of the zone of interest. The different boundary condition also play a part, where the top row used PEC, and the other two rows PMC.

Figure~\ref{fig:results-marlim_survey} shows a zoom-in to the survey domain, the domain of interest, which shows how the different codes discretized the zones of interest.
%
\plot*{results-marlim_survey}{width=\textwidth}{Meshes and broadside $E_x$ and $E_y$ fields in the left, middle, and right column, respectively, for the rectilinear, octree, and tetrahedral meshes in the top, middle, and bottom row, respectively (3$\times$ vertical exaggeration).}
%
The most striking point of these mesh comparison plots is that, while the overall fields are similar and particularly the responses at receiver locations are the same, the fields can vary quite a lot in different parts of the model. E.g., two kilometres below the seafloor there are already significant changes visible, purely due to meshing. This is an important insight that has to be taken into account when interpreting inversion results.

The required runtime and memory to compute the shown MR3D responses for the four codes are listed in Table~\ref{tbl:comp-marlim}. The memory requirement varies from $0.5-230\,$GiB, and the runtime from roughly 7\,min to 20\,min, where the codes were run on very different machines from servers to supercomputers (see Table \ref{tbl:machines}) and use between 1 and 96 processes in parallel.
%
\tabl[bthp]{comp-marlim}{Comparison of number of processes, runtime, and memory, as well as the degree of freedom of the discretization used by the different codes for the MR3D model.}{
  \centering
\begin{tabular}{lrS[table-format=6.0]S[table-format=4.1]S[table-format=8.0]}
  \toprule
  %
       & \#Procs & {Runtime (s)} & {Memory (GiB)}   & {\#dof} \\
  \midrule
  %
  \custem & 64 &  872 & 230.1 & 1918106 \\ % dof: 3836212 / 2 (Real+Imag -> Complex)
  \emg3d  &  1 & 1200 &   0.5 & 5998992 \\
  \petgem & 96 &  524 & 175.4 & 1918106 \\
  \simpeg &  4 &  422 &  12.8 &  720146 \\ % nprocs: 1*4 threads
  %
  \bottomrule
\end{tabular}}%
%

The actual discretized resistivity model of the three mesh types are shown in Figure~\ref{fig:results-marlim_allmodels}.
%
\plot*{results-marlim_allmodels}{width=\textwidth}{The resistivity model as obtained from the three different discretizations using rectilinear, octree, and tetrahedral meshes (3$\times$ vertical exaggeration).}
%
The models where obtained using volume averaging (\emg3d, \simpeg) and linear interpolation (\custem, \petgem), in both cases on $\log_{10}{\rho}$ values. The actual representation of the original resistivities in the tetrahedra mesh is not as smooth as in the other meshes since this would require significantly smaller tetrahedra and more dof on top of the inclusion of subsurface layer constraints. However, note that the combination of slicing through tetrahedra and the vertical exaggeration lead to a very distorted view of the tetrahedral mesh which is barely representative for the overall averaged approximation of resistivities. The actual responses at the receivers (field interpolation) are obtained by spline interpolation (\emg3d), linear interpolation (\simpeg), and through the second-order basis functions (\custem, \petgem).

@@

%% 1. Results: good agreement %%
Overall, we observed an excellent match between the solutions. The amplitudes of the layered model have mostly a relative error of less than $1-2\,$\%. Looking individually at the real or imaginary part results in much higher error values, but those are related to the zero crossings of the responses. The cross-validation of the results for the block model yields a similar picture, with NRMSD values within a few percents. For the large, complex MR3D model it is a bit different. First it can be noted that the errors seem to be higher for higher frequencies. For the lowest frequency, 0.125\,Hz, the NRMSD is a few percent for inline and broadside $E_x$ components, and up to 10\,\% for the broadside $E_y$ component. For higher frequencies these numbers rise to the order of $\pm10\,$\%. The reason for the higher misfit for MR3D than for the layered or block model lies mostly in the discretization, boundary effects, and interpolation strategy. While one might argue that $5-10\,$\% NRMSD is a lot, we argue that this is probably as good as it gets, given that the code with which the responses we compare to were computed is not open-source. This has been shown within the process of this study. Our first attempt were based on the originally published, fine MR3D model. This model is very detailed, but its extent is too small in x- and y-directions for CSEM computations, and each of our codes did its own extrapolation. However, none of the codes were able to reproduce the published results. The reason is that there are many ways how you can extrapolate a model. We reached out to the authors of MR3D, and they kindly agreed to also published the upscaled and extended computational model. It is only with this model that we were able to get comparable results.

@@

All files to rerun the different models with the four codes and reproduce the shown results are available at Zenodo (DOI) \emph{with the final submission}. For the review process we refer to the github page \href{https://github.com/prisae/3d-csem-open-source-landscape}{github.com/prisae/3d-csem-open-source-landscape}.

>>>>

% Additional packages
\usepackage[UKenglish]{babel}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{amssymb, amsmath, amsfonts}
\usepackage{siunitx}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{xspace}
\usepackage{flafter}
\usepackage[pdftex, final, allcolors=blue, colorlinks=true]{hyperref}
\frenchspacing

@@

\begin{abstract}
Large-scale modelling of three-dimensional controlled-source electromagnetic (CSEM) surveys used to be feasible only for large companies and research consortia. This has changed over the last few years, and today there exists a selection of different open-source codes. Using four different codes in the Python ecosystem, we perform simulations for increasingly complex models. We first verify the computed fields with semi-analytical solutions for a simple layered model. Then we validate the responses of a more complex block model by comparing results obtained from each code. Finally we compare the responses of a real world model with results from the industry. These validations show that the open-source codes are able to compute comparable CSEM responses for challenging, large-scale models. Our comparison includes finite-element and finite-volume codes using structured rectilinear and octree meshes as well as unstructured tetrahedral meshes. Accurate responses can be obtained independently of the chosen method and the chosen mesh type. The runtime and memory requirements vary greatly based on the choice of iterative or direct solvers. However, we have found that much more time was spent on designing the mesh and setting up the simulations than running the actual computation. The challenging task is, irrespective of the chosen code, to appropriately discretize the model. We provide three models, each with their corresponding discretization and responses of four codes, which can be used for validation of new and existing codes. The collaboration of four code maintainers trying to achieve the same task brought in the end all four codes a significant step further. This includes improved meshing and interpolation capabilities, resulting in shorter runtimes for the same accuracy. We hope that these results may be useful for the CSEM community at large and that we can build over time a suite of benchmarks that will help to increase the confidence in existing and new 3D CSEM codes.
%
% https://academic.oup.com/DocumentLibrary/GJI/gji%20keyword%20list%20updated2017.pdf
\\[1cm] \textbf{Keywords}: Controlled source electromagnetics (CSEM), Numerical modelling, Electrical properties
\end{abstract}

@@

The open-source landscape in the field of 3D CSEM modelling changed in the last five years quite dramatically, and here we introduce and compare four recent open-source projects. All presented codes are in the Python ecosystem and use either the FV method on structured grids or the FE method on unstructured tetrahedral meshes. Having different codes which use different methods and different meshes is ideal to address the topic of validation. Analytical and semi-analytical solutions only exist for simple halfspace or layered-Earth models, which served mostly to verify new codes. The only objective possibility to ensure the accuracy of solutions beyond these simple models is by comparing results from different modellers. If different discretizations and implementations of Maxwell's equations yield the same result, it gives confidence in their accuracy. \cite{GJI.13.Miensopust} presents a review of two workshops dealing with the validation of MT forward and inversion codes, but we are not aware of any comparable comparison study or benchmark suite for CSEM data.

@@

We briefly present our codes in this section. It is, however, beyond the scope of this article to go into every detail of the different modellers, and we refer to their documentations for more details. An overview comparison of the codes is given in Table~\ref{tbl:codecomparison}. All four have in common that they solve Maxwell's equation in its differential form (DE) under the quasi-static or diffusive approximation, hence neglecting displacement currents, which is a common approximation for the low frequency range usually applied in CSEM. The machines on which the different codes were run are listed in Table~\ref{tbl:machines} together with the responsible operator. A few clarifying words on abbreviations and definitions: For the boundary conditions (BC) in the comparison table we use the abbreviations \emph{PEC} and \emph{PMC}, which stand for perfect electric conductor
% We could add maths here; but we would need to introduce all the symbols. For now, we leave it out.
% ($\vec{n}\times\vec{E} = 0; \vec{n}\cdot\vec{H} = 0$)
and perfect magnetic conductor%
% ($\vec{n}\cdot\vec{E} = 0; \vec{n}\times\vec{H} = 0$)
, respectively. Another abbreviation used in the tables is \emph{dof} for degree of freedom, which is equivalent to the size of the SLE we are solving. Finally, we use \emph{runtime} as the wall time or elapsed real-world time from start to end of solving the SLE; pre- and post-processing (e.g., mesh generation) is not measured. Note that the actual computation time or CPU time is therefore much higher than the reported runtime for the codes that run in parallel, with an upper limit of \#Procs$\times$Runtime. Memory refers to the maximum memory increase at any point of solving the SLE.

\tabl[bthp]{codecomparison}{Comparison of the four codes under consideration. Note that \emg3d is a solver on its own, while the other codes implement third-package solvers such as \texttt{PETSc} \citep{Preprint.Abhyankar}, \texttt{MUMPS} \citep{SIAM.01.Amestoy}, or \texttt{PARDISO} \citep{FGCS.04.Schenk}.}{
  \centering
  \footnotesize
\begin{tabularx}{\linewidth}{lXXXX}
  \toprule
  %
  & \custem & \emg3d & \petgem & \simpeg  \\
  \midrule
  %
  Website & \href{https://custem.rtfd.io}{custem.rtfd.io}
          & \href{https://empymod.github.io}{empymod.github.io}
          & \href{http://petgem.bsc.es}{petgem.bsc.es}
          & \href{https://docs.simpeg.xyz}{simpeg.xyz} \\
  License & GPL-3.0 & Apache-2.0 & BSD-3-Clause & MIT \\
  Installation & \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{pip}
               & \texttt{pip}; \texttt{conda} \\
  Comp. Dom. & frequency \& time & frequency & frequency & frequency \& time \\
  Method & FE & FV & FE & FV \\
  Mesh & tetrahedral & rectilinear & tetrahedral & recti-/curvilinear, octree \\
  BC & PEC; PMC & PEC & PEC & PEC; PMC \\
  Solver & \texttt{MUMPS} & \emg3d & \texttt{PETSc}; \texttt{MUMPS} &
           \texttt{PARDISO}; \texttt{MUMPS} \\
  %
  \bottomrule
\end{tabularx}}%
%

@@

I few interesting points stand out:
\begin{enumerate}
  \item The error levels between the different codes are overall comparable; the NRMSD is generally below a few percent, increasing towards larger offsets.
  \item The NRMSD between the FV codes \emg3d and \simpeg is comparably low except at the boundary. The generally low NRMSD is because these codes use the same rectilinear mesh for modelling. The increasing NRMSD towards the boundary is because \emg3d uses the PEC boundary condition, while \simpeg uses the PMC boundary condition in this example.
  \item The FE codes \custem and \petgem show similar boundary effects, but are influenced by the anomaly response (higher errors on right side and now unsymmetrical error distribution).
  \item The agreement within the same differential equation method (FE or FV) is better than across methods. 
\end{enumerate}

@@

The responses for all six frequencies and all three electric components are shown in Figure~\ref{fig:results-marlim-responses}, both inline and broadside. The published responses are shown in colour with markers for the frequency, and the responses from our codes are shown beneath in grey colours in this overview plot.
%
\plot*{results-marlim-responses}{width=\textwidth}{MR3D comparison between our codes (grey lines) and the published data (coloured lines). The grey lines under the coloured lines are not visible in most areas, meaning that they are very similar. However, there are three notable zones, (1) to (3), which are explained further in the text.}
%
Two CSEM data-sets were published with MR3D, one containing clean responses, and one responses where realistic noise was added. For the comparison we use the clean data, but we indicate the chosen noise level by the horizontal line at \num{2e-15}\,V/m. It can be seen that the grey lines are not visible for most parts, which means that the data agree well. However, there are a few notable points where this is not the case, and they are marked with numbers in the figure.
\begin{enumerate}
  \item There are some noticeable differences for positive offsets of the broadside $E_y$ and $E_z$ components, which are likely related to the bathymetry. The reason why we cannot reproduce the published results exactly is that the code is not open-source, and we do not exactly know what it does internally. See the discussion around Figure~\ref{fig:results-marlim_2published}.
  \item The three highest frequencies become noisy at large offsets for all of our codes (well below any real-world noise level). The reason why the published responses are not noisy is probably because they were computed in the time domain, and transformed to the frequency domain. This explains their smooth behaviour, without saying anything about their accuracy.
  \item The responses for the $E_y$ component of the inline acquisition line do not agree at all. In the 1D case, the inline $E_y$ component would be zero, and the only response we can measure here comes from 3D effects. These responses are very low, roughly two orders of magnitude lower than the $E_x$ component. Not the published nor our responses are stable, and any of the responses is as bad as the other. Tiniest differences in meshing or different interpolation algorithms will have a huge effect here.
\end{enumerate}

It is important to note that the published results nor any of our results is the \emph{thruth}, but the closer they agree with each other the more certain we can be of the outcomes. Figure~\ref{fig:results-marlim_2published} shows the NRMSDs between the published results and our codes for the three strongest components, the inline $E_x$ and the broadside $E_x$ and $E_y$ components, for three frequencies.
%
\plot*{results-marlim_2published}{width=\textwidth}{NRMSDs for all four codes in comparison with the published data, for the inline $E_x$ field and the broadside $E_x$ and $E_y$ fields in the left, middle, and right column, respectively. Shown are the three frequencies 0.125\,Hz, 0.5\,Hz, and 1.0\,Hz in the top, middle, and bottom row, respectively.}
%
A few conclusions can be drawn from this figure: In general the NRMSD is roughly 10\,\% or less. The structured meshes have in some cases a smaller NRMSD, e.g., for negative offsets in the inline and broadside $E_x$ responses, and in other cases the unstructured tetrahedral meshes, e.g., for positive offsets for all broadside $E_y$ responses, and often they have a comparable NRMSD. Some NRMSDs have an interesting step-pattern, which is more pronounced for the structured meshes than for the unstructured meshes; this is related to the annotated point (3) in Figure~\ref{fig:model-marlim}. The origin of this zigg-zagg pattern is most likely the bathymetry. We suspect that the code of the published data did probably use the bathymetry information in addition to the resistivity cube. This would explain why the codes using the tetrahedra mesh, which uses the bathymetry as well, does not show this feature.

The importance of the meshing to the result can be seen in Figure~\ref{fig:results-marlim_2ours}, where we compute the NRMSDs between our codes.
%
\plot*{results-marlim_2ours}{width=\textwidth}{NRMSDs, just as in Figure~\ref{fig:results-marlim_2published}, but comparing some of the four codes with each other. \custem and \petgem produce very similar results, which is due to the fact that they use the same mesh. This also shows that the biggest impact in 3D forward modelling comes from the discretization.}
%
The most obvious result in that figure is that \custem and \petgem produce results that are almost identical, their NRMSD is generally below 0.03\,\%. The reason is simple: \custem and \petgem use the exactly same mesh for the computation of this example, the only difference is therefore the solver. This exemplifies that the biggest difference between different codes is the discretization; the solver is important when it comes to runtime and memory, but not for the accuracy of the result. As these two are so similar we compare \emg3d and \simpeg only to \custem in Figure~\ref{fig:results-marlim_2ours}, as the comparisons to \petgem would look the same. But in general the NRMSDs between our codes look similar as in the comparison with the published results, and the principal source of the difference must be in the different discretizations.

The actual differences in meshing is shown in the next two figures. Figure~\ref{fig:results-marlim_big} shows on the top row the rectilinear mesh of \emg3d, on the middle row the octree mesh of \simpeg, and on the bottom row the tetrahedral mesh of \custem and \petgem. The actual mesh and the broadside $|E_x|$ and $|E_y|$ fields are shown in the left, middle, and right columns, respectively. The extent of the plot includes the entire computational domain for the tetrahedral mesh (in x and z direction); the rectilinear and octree mesh extents are larger than what is shown (rectilinear: $x=327$ to 454\,km, $z=-57$ to 66\,km; octree: $x=288$ to 492\,km, $z=-123$ to 41\,km).
%
\plot*{results-marlim_big}{width=\textwidth}{Meshes and broadside $E_x$ and $E_y$ fields in the left, middle, and right column, respectively, for the rectilinear, octree, and tetrahedral meshes in the top, middle, and bottom row, respectively. The shown dimensions shows the entire computational domain for the tetrahedra mesh, whereas the entire rectilinear and octree meshes are bigger.}
%
This figure is interesting for various points, and the differences between the three meshes are very obvious. Particularly visible is the advantage of octree and tetrahedra to create large cells outside of the region of interest to push away the boundary, whereas a rectilinear grid, even though stretched, has to compute many cells far away of the zone of interest. The different boundary condition also play a part, where the top row used PEC, and the other two rows PMC.

Figure~\ref{fig:results-marlim_survey} shows a zoom-in to the survey domain, the domain of interest, which shows how the different codes discretized the zones of interest.
%
\plot*{results-marlim_survey}{width=\textwidth}{Meshes and broadside $E_x$ and $E_y$ fields in the left, middle, and right column, respectively, for the rectilinear, octree, and tetrahedral meshes in the top, middle, and bottom row, respectively (3$\times$ vertical exaggeration).}
%
The most striking point of these mesh comparison plots is that, while the overall fields are similar and the responses at receiver locations agree well, the fields can vary quite a lot in different parts of the model. Two kilometres below the seafloor, as an example, there are already significant changes visible, purely due to meshing. This is an important insight that has to be taken into account when interpreting inversion results.

The required runtime and memory to compute the shown MR3D responses for the four codes are listed in Table~\ref{tbl:comp-marlim}. The memory requirement varies from $0.5-230\,$GiB, and the runtime from roughly 7\,min to 20\,min, where the codes were run on very different machines from servers to supercomputers (see Table \ref{tbl:machines}) and use between 1 and 96 processes in parallel.
%
\tabl[bthp]{comp-marlim}{Comparison of number of processes, runtime, and memory, as well as the degree of freedom of the discretization used by the different codes for the MR3D model.}{
  \centering
\begin{tabular}{lrS[table-format=6.0]S[table-format=4.1]S[table-format=8.0]}
  \toprule
  %
       & \#Procs & {Runtime (s)} & {Memory (GiB)}   & {\#dof} \\
  \midrule
  %
  \custem & 64 &  872 & 230.1 & 1918106 \\ % dof: 3836212 / 2 (Real+Imag -> Complex)
  \emg3d  &  1 & 1254 &   0.6 & 5998992 \\
  \petgem & 96 &  524 & 175.4 & 1918106 \\
  \simpeg &  4 &  422 &  12.8 &  720146 \\ % nprocs: 1*4 threads
  %
  \bottomrule
\end{tabular}}%
%

The actual discretized resistivity model of the three mesh types are shown in Figure~\ref{fig:results-marlim_allmodels}.
%
\plot*{results-marlim_allmodels}{width=\textwidth}{The resistivity model as obtained from the three different discretizations using rectilinear, octree, and tetrahedral meshes (3$\times$ vertical exaggeration).}
%
The models where obtained using volume averaging (\emg3d, \simpeg) and linear interpolation (\custem, \petgem), in both cases on $\log_{10}(\rho)$ values. The actual representation of the original resistivities in the tetrahedra mesh is not as smooth as in the other meshes since this would require significantly smaller tetrahedra and more dof on top of the inclusion of subsurface layer constraints. However, note that the combination of slicing through tetrahedra and the vertical exaggeration leads to a very distorted view of the tetrahedral mesh which is barely representative for the overall averaged approximation of resistivities. The actual responses at the receivers (field interpolation) are obtained by spline interpolation (\emg3d), linear interpolation (\simpeg), and through the second-order basis functions (\custem, \petgem). It is worth mentioning as a final note that we used frequency independent meshes in this study. Creating frequency dependent meshes could potentially improve both the result and the runtime.

@@

%% 1. Results: good agreement %%
Overall, we observed an excellent match between the solutions. The amplitudes of the layered model have mostly a relative error of less than 1--2\,\%. Looking individually at the real or imaginary part results in much higher error values, but those are related to the zero crossings of the responses. The cross-validation of the results for the block model yields a similar picture, with NRMSD values within a few percents. For the large, complex MR3D model it is a bit different. First it can be noted that the misfits are higher for higher frequencies. For the lowest frequency, 0.125\,Hz, the NRMSD is a few percent for inline and broadside $E_x$ components, and up to 10\,\% for the broadside $E_y$ component. For higher frequencies these numbers rise to the order of $\pm10\,$\%. The reason for the higher misfit for MR3D than for the layered or block model lies mostly in the discretization, boundary effects, and interpolation strategy. While one might argue that 5--10\,\% NRMSD is a lot, we argue that this is probably as good as it gets, given that the code with which the responses we compare to were computed is not open-source. This has been shown within the process of this study. Our first attempt were based on the originally published, fine MR3D model. This model is very detailed, but its extent is too small in x- and y-directions for CSEM computations, and each of our codes did its own extrapolation. However, none of the codes were able to reproduce the published results. The reason is that there are many ways how you can extrapolate a model. We reached out to the authors of MR3D, and they kindly agreed to also published the upscaled and extended computational model. It is only with this model that we were able to get comparable results.

@@

All files to rerun the different models with the four codes and reproduce the shown results are available at Zenodo (DOI) \emph{with the final submission}. For the review process we refer to the github page \href{https://github.com/swung-research/3d-csem-open-source-landscape}{github.com/swung-research/3d-csem-open-source-landscape}.
